---
title: "GroupProject1"
author: "Us"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(car)
```

```{r}
premier <- read_csv("premier1819.csv") %>%
  filter(position %in% c("Forward")) %>%
  select(appearances = appearances_overall, goals = goals_overall)
summary(premier)
```

```{r}
premier_plot <- ggplot(data = premier, mapping = aes(x = appearances, y = goals)) +
  geom_point()
cor(premier)
```

```{r}
premier <- premier %>%
  mutate(goals = goals + 1)
bc <- boxCox(premier$goals ~ premier$appearances)  # plot curve
bc$x[which.max(bc$y)]  # pull out the "best" lambda value

premier_trans <- premier %>%
  mutate(goals = 1/sqrt(goals))
(premier_trans_plot <- ggplot(data = premier_trans, mapping = aes(x = appearances, y = goals)) +
  geom_point())
cor(premier)

premier_lm <- lm(goals ~ appearances, data = premier_trans)
premier$residuals <- premier_lm$residuals
```
# Check the model assumption

## L) Linear
```{r}
# Scatterplot
premier_trans_plot

# Resids vs fitted
premier_residfit <- autoplot(premier_lm, which = 1, ncol = 1) + theme_classic()
```

## I) Independent
The residuals are independent: Considering there can only be so many forwards on a given team, it would be very difficult to obtain the data for two forwards at the same time. The only case where two forwards would be in the same place at the same time is when they were on opposing teams, so they would have little to no effect on each other.

## N) Normal
```{r}
# Q-Q Plot
(premier_QQ <- autoplot(premier_lm, which = 2, ncol = 1, nrow = 1))

# Histogram
premier_hist <- ggplot(data = premier, mapping = aes(x = residuals)) + 
  # when using this code for future data sets, make sure to change the binwidth: 
  geom_histogram(mapping = aes(y = ..density..), binwidth = 0.012) +
  # stat_function() overlays the red normal curve on the histogram
  stat_function(fun = dnorm, 
                color = "red", 
                size = 2,
                args = list(mean = mean(premier$residuals), 
                            sd = sd(premier$residuals)))

premier_hist

# Box-plot
ggplot(data = premier, mapping = aes(y = residuals)) + geom_boxplot()
```


## E) Equal
```{r}
# Residuals vs Fitted Values
premier_residfit

# Brown-Forsythe Test
grp <- as.factor(c(rep("lower", floor(dim(premier)[1] / 2)), 
                   rep("upper", ceiling(dim(premier)[1] / 2))))
leveneTest(arrange(premier, appearances)$residuals ~ grp, center = median)
```

## A) All
```{r}
#Q-Q
premier_QQ

# Cook's Distance
premier_trans <- premier_trans %>%
  mutate(cooksd = cooks.distance(premier_lm))
top4cd <- as.numeric(names(sort(premier_trans$cooksd, decreasing = TRUE)[1:4]))

ggplot() +
  geom_point(data = premier_trans,
             mapping = aes(x = as.numeric(rownames(premier_trans)), 
                           y = cooksd)) +
  geom_text(mapping = aes(x = top4cd,
                          y = premier_trans$cooksd[top4cd],
                          label = top4cd)) +
  theme_bw() +
  ylab("Cook's Distance") +
  xlab("Observation Number") +
  geom_hline(mapping = aes(yintercept = 4 / length(premier_trans$cooksd)),
             color = "red", linetype = "dashed") +
  theme(aspect.ratio = 1)

# DFBETAS
premier_trans <- premier_trans %>%
  mutate(dfbetas_appearances = dfbetas(premier_lm)[, "appearances"])
names(premier_trans$dfbetas_appearances) <- 1:nrow(premier_trans)
top3dfbeta <- as.numeric(names(
  sort(abs(premier_trans$dfbetas_appearances), decreasing = TRUE)[1:3]
))

# Plot the DFBETAS against the observation number
ggplot() + 
  geom_point(data = premier_trans,
             mapping = aes(x = as.numeric(rownames(premier_trans)), 
                           y = abs(dfbetas_appearances))) +
  geom_text(mapping = aes(x = top3dfbeta,
                          y = abs(premier_trans$dfbetas_appearances[top3dfbeta]),
                          label = top3dfbeta)) +
  theme_bw() +
  ylab("Absolute Value of DFBETAS for Runoff") +
  xlab("Observation Number") +
  geom_hline(mapping = aes(yintercept = 2 / sqrt(length(premier_trans$dfbetas_appearances))),
             color = "red", linetype = "dashed") +
  theme(aspect.ratio = 1)

#Resid vs Fitted
premier_residfit
```

## R) Required
There aren’t many other additional predictor variables that are likely to help in predicting the response variable of goals scored. The only thing that may affect the number of goals scored per forward would be the weather in a given match, but it varies pretty randomly and would have an equal effect on all forwards, so there isn’t much extra variability that isn’t accounted for.