---
title: "Homework 5"
subtitle: <center> <h1>Multiple Linear Regression Variable Selection Methods</h1> </center>
author: <center> Joshua Carpenter <center>
output: html_document
---

<style type="text/css">
h1.title {
font-size: 40px;
text-align: center;
}
</style>

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)  # plot lm objects using ggplot instead of base R
library(car)  # needed for added-variable plots and dfbetas and dffits
library(corrplot)  # colored correlation matrix
library(gridExtra)  # NEW PACKAGE for putting multiple ggplots in one plot
set.seed(6865)
```

### Data and Description

**For this assignment, we are revisiting the data set used in Homework 4. I think it would be very beneficial for you to review your Homework 4 before starting this one.**

*Note that you do not need to properly format the axis limits in your plots for this assignment (to save time). You should, however, still make the plots square in shape.*

Measuring body fat is not simple. One method requires submerging the body underwater in a tank and measuring the increase in water level. A simpler method for estimating body fat would be preferred. In order to develop such a method, researchers recorded age (years), weight (pounds), height (inches), and three body circumference measurements (around the neck, chest, and abdominal (all in centimeters)) for 252 men. Each man’s percentage of body fat was accurately estimated by an underwater weighing technique (the variable brozek is the percentage of body fat). The hope is to be able to use this data to create a model that will accurately predict body fat percentage, by using just the basic variables recorded, without having to use the tank submerging method. 

The data can be found in the BodyFat data set on Canvas. Download BodyFat.txt, and put it in the same folder as this R Markdown file.

#### 0. Replace the text "< PUT YOUR NAME HERE >" (above next to "author:") with your full name.

#### 0b. Make sure to set your seed since some of the functions randomly split your data (use `set.seed` in the setup code chunk above)!

#### 1. Read in the data set, and call the data frame "bodyfat_orig". Print a summary of the data and make sure the data makes sense. 

```{r}
bodyfat_orig <- read.csv("bodyfat.txt", sep = " ", row.names = 1)
(bodyfat_orig <- tibble(bodyfat))
summary(bodyfat_orig)
```

#### 2. Refer back to your Homework 4. In that assignment, you fit this multiple linear regression model: for each of the multiple linear regression assumptions listed below, state if they were met or not met.

1. The X’s vs Y are linear:   Met
2. The residuals are normally distributed and centered at zero:   Met
3. The residuals are homoscedastic:   Met
4. The model describes all observations (i.e., there are no influential points):   Not met
5. No multicollinearity:   Not met

#### 3. There is one clear influential point in the data set. Create a new variable called "bodyfat" that contains the bodyfat_orig data set with the influential point removed. Use the bodyfat data set throughout the rest of the assignment.

```{r, fig.align='center'}
bodyfat <- bodyfat_orig %>%
  slice(-39)
```


### You should have discovered, from Homework 4, that there is a multicollinearity problem. The goal of this assignment is to continue this analysis by identifying variables to potentially remove from the model to resolve the multicollinearity issues. 




#### 4. Briefly explain why multicollinearity is a problem for multiple linear regression by identifying some (at least two) of the consequences of multicollinearity.

One of the consequences of multicollinearity is inflated variance which leads to wide confidence intervals. Another consequence is potentially inaccurate coefficients and hypothesis tests; variable can appear significant that aren't and vise versa and coefficients can have the opposite sign that they should.

#### 5. Briefly explain the similarities and differences between the following methods: best subset, forward, backward, and sequential replacement. Do not just copy the algorithm from the class notes - use your own words to explain what these methods are doing.

The best subset is the most thorough and preferable method. It compares all possible models and selects the best one. Forward selection is basically useless, but is the basis for the other two. It starts with an intercept model and adds variables that improve the model. Backward does the opposite; it starts with all variables and removes the ones that do not improve the model. Sequential starts with an intercept only model, adds variables one by one that improve the model and with each step also removes variables that don't improve the model. When best subsets is not possible, sequential replacement is probably the best bet.

#### 6. Briefly explain how shrinkage methods work (variance-bias tradeoff). Specifically, how can some of these methods be considered variable selection procedures?

Shrinkage methods introduce a little bit of bias in order to greatly reduce variance. Sometimes we are willing to make that trade off in order to get significant results.

#### 7. Briefly explain the similarities/difference between ridge regression, LASSO, and elastic net.

All three methods shrink coefficients 

#### 8. Remember, when coding these methods, the response variable must be the last column in the data set for the `bestglm` function to work. Switch the order of the columns in the data set so that brozek is last.

```{r, fig.align='center'}
# your code here
```

#### 9. Apply the best subsets variable selection procedure to this data set. You may choose which metric you would like to use (ex: AIC, BIC, PMSE). Output a summary of the "best" model.

```{r, fig.align='center'}
# your code here
```

#### 10. Apply the forward selection procedure to this data set. You may choose which metric you would like to use (ex: AIC, BIC, PMSE). Output a summary of the "best" model.

```{r, fig.align='center'}
# your code here
```

#### 11. Apply the backward selection procedure to this data set. You may choose which metric you would like to use (ex: AIC, BIC, PMSE). Output a summary of the "best" model.

```{r, fig.align='center'}
# your code here
```

#### 12. Apply the sequential replacement selection procedure to this data set. You may choose which metric you would like to use (ex: AIC, BIC, PMSE). Output a summary of the "best" model.

```{r, fig.align='center'}
# your code here
```

#### 13. Apply LASSO to this data set using the MSE metric. Output the coefficient values corresponding to the 1 standard error rule (do not output any plots).

```{r, fig.align='center'}
# your code here
```

#### 14. Apply Elastic Net to this data set using the MSE metric. Output the coefficient values corresponding to the 1 standard error rule (do not output any plots).

```{r, fig.align='center'}
# your code here
```  

#### 15. Fill in the table below with "X"s (like the one at the end of the course notes: a row for each variable, a column for each variable selection method, an "X" in a cell means the variable was included for that variable selection method).

Variable     | Best Subset | Forward | Backward | Sequential Replacement | LASSO  | Elastic Net
------------ | ----------- | ------- | -------- | ---------------------- | ------ | -----------
age          |             |         |          |                        |        |
weight       |             |         |          |                        |        |
height       |             |         |          |                        |        |
neck         |             |         |          |                        |        |
chest        |             |         |          |                        |        |
abdom        |             |         |          |                        |        |




#### 16. Now that you have seen the various results from the different methods, pick a subset of variables that you will include in the model. Which variables do you choose to include in the model? Why?

< your response here >

#### 17. Create the multiple linear regression model with the variables you listed in the previous question (alternatively, you can call the best model using $BestModel). Print a summary of the results.

```{r, fig.align='center'}
# your code here
```





### Now that you have chosen a model, the next several questions ask you to check all of the model assumptions. For each assumption, (1) perform appropriate diagnostics to determine if the assumption is violated, and (2) explain whether or not you think the assumption is violated and why you think that. **Note: you can copy (then modify) a lot of your code from Homework 4 to answer these questions.**





#### 18. (L) The Xs vs Y are linear (use all four diagnostics)

```{r, fig.align='center'}
# your code here
```

< your response here >

#### 19. (I) The residuals are independent (no diagnostic tools - just think about how the data was collected and briefly write your thoughts)

< your response here >

#### 20. (N) The residuals are normally distributed and centered at zero (use all four diagnostics)

```{r, fig.align='center'}
# your code here
```

< your response here >

#### 21. (E) The residuals have equal/constant variance across all values of X (use the one diagnostic tool)

```{r}
# your code here
```

< your response here >

#### 22. (A) The model describes all observations (i.e., there are no influential points) (use at least four diagnostic tools)

```{r, fig.align='center'}
# your code here
```

< your response here >

#### 23. (R) Additional predictor variables are not required (no diagnostic tools - just think about the variables you have and if there are other variables you think would help predict the response)

< your response here >

#### 24. No multicollinearity  (use all three diagnostics)

```{r, fig.align='center'}
# your code here
```

< your response here >

#### 25. Given the results from your model assumption checking, what would you do next to continue this analysis?

< your response here >

#### 26. Briefly summarize what you learned, personally, from this analysis about the statistics, model fitting process, etc.

< your response here >

#### 27. Briefly summarize what you learned from this analysis *to a non-statistician*. Write a few sentences about (1) the purpose of this data set and analysis and (2) what you learned about this data set from your analysis. Write your response as if you were addressing a business manager (avoid using statistics jargon) and just provide the main take-aways.

< your response here >